---
title: "CodeBook"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load(".Rdata")
```

## Overview

Both of train and test data are stored in a data frame called **activities**.

```{r}
summary(activities) #tidy version
```


## The Data


<table>


<tr><th>File Name</th><th>Description</th></tr>

<tr><td valign=top>`./UCI HAR Dataset/activity_labels.txt`</td><td>Vavriable storing the whole data</td></tr>

<tr><td valign=top>`./UCI HAR Dataset/features.txt`</td><td> Variable storing activity labels in `./UCI HAR Dataset/activity_labels.txt`</td></tr>

<tr><td valign=top>`./UCI HAR Dataset/test/subject_test.txt`</td><td> intervening variable storing mean values of all observations</td></tr>

<tr><td valign=top>`./UCI HAR Dataset/test/X_test.txt`</td><td> intervening variable storing standard deviation values of all observations </td></tr>

<tr><td valign=top>`./UCI HAR Dataset/test/y_test.txt`</td><td> variable storing data of `./UCI HAR Dataset/features.txt`</td></tr>

<tr><td valign=top>`./UCI HAR Dataset/train/subject_train.txt`</td><td> variable storing activity labels of test subject in `./UCI HAR Dataset/test/y_test.txt`  </td></tr>

<tr><td valign=top>`./UCI HAR Dataset/train/X_train.txt`</td><td> variable storing activity labels of test subject in `./UCI HAR Dataset/test/y_test.txt`  </td></tr>

<tr><td valign=top>`./UCI HAR Dataset/train/y_train.txt`</td><td> variable storing activity labels of test subject in `./UCI HAR Dataset/test/y_test.txt`  </td></tr>



</table>


## The Variables

<table>


<tr><th>Varible Name</th><th>Description</th></tr>

<tr><td valign=top>activities</td><td>Vavriable storing the whole data</td></tr>

<tr><td valign=top>activity_labels</td><td> Variable storing activity labels in `./UCI HAR Dataset/activity_labels.txt`</td></tr>

<tr><td valign=top>data.mean</td><td> intervening variable storing mean values of all observations</td></tr>

<tr><td valign=top>data.std</td><td> intervening variable storing standard deviation values of all observations </td></tr>

<tr><td valign=top>features</td><td> variable storing data of `./UCI HAR Dataset/features.txt`</td></tr>

<tr><td valign=top>label_test</td><td> variable storing activity labels of test subject in `./UCI HAR Dataset/test/y_test.txt`  </td></tr>

<tr><td valign=top>label_train</td><td>variable storing activity labels of train subject in `./UCI HAR Dataset/train/y_train.txt` </td></tr>

<tr><td valign=top>set_test</td><td> variable storing signal data for test in `./UCI HAR Dataset/test/X_test.txt`</td></tr>

<tr><td valign=top>set_train</td><td>variable storing signal data for train in `./UCI HAR Dataset/train/X_train.txt`</td></tr>

<tr><td valign=top>subject_test</td><td> variable storing test subject id of `./UCI HAR Dataset/test/subject_test.txt` </td></tr>

<tr><td valign=top>subject_train</td><td>variable storing train subject id of `./UCI HAR Dataset/train/subject_train.txt` </td></tr>

<tr><td valign=top>test</td><td> intervening variable containing all test data</td></tr>

<tr><td valign=top>train</td><td> intervening variable containing all train data</td></tr>


</table>



## The Key Pacakges and Functions

<table>


<tr><th>Name</th><th>Description</th></tr>


<tr><td valign=top>`dplyr::arrange()`</td><td> `arrange()` order the rows of a data frame rows by the values of selected columns.</td></tr>

<tr><td valign=top>`dplyr::as_tibble()`</td><td> `as_tibble()` turns an existing object, such as a data frame or matrix, into a so-called tibble, a data frame with class tbl_df. </td></tr>

<tr><td valign=top>`dplyr::bind_cols()`</td><td> This is an efficient implementation of the common pattern of do.call(rbind, dfs) or do.call(cbind, dfs) for binding many data frames into one. </td></tr>

<tr><td valign=top>`dplyr::bind_rows()`</td><td> This is an efficient implementation of the common pattern of do.call(rbind, dfs) or do.call(cbind, dfs) for binding many data frames into one.
</td></tr>

<tr><td valign=top>`dplyr::mutate()`</td><td>`mutate()` adds new variables and preserves existing ones; `transmute()` adds new variables and drops existing ones. New variables overwrite existing variables of the same name. Variables can be removed by setting their value to NULL.</td></tr>

<tr><td valign=top>`dplyr::select()`</td><td> Select (and optionally rename) variables in a data frame, using a concise mini-language that makes it easy to refer to variables based on their name (e.g. a:f selects all columns from a on the left to f on the right). You can also use predicate functions like is.numeric to select variables based on their properties.

</td></tr>

<tr><td valign=top>`tidyr::pivot_longer()`</td><td> `pivot_longer()` "lengthens" data, increasing the number of rows and decreasing the number of columns. The inverse transformation is `pivot_wider()`.

</td></tr>

<tr><td valign=top>`tidyr::pivot_wider()`</td><td> `pivot_wider()` "widens" data, increasing the number of columns and decreasing the number of rows. The inverse transformation is `pivot_longer()`. </td></tr>


</table>



