---
title: "README"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)


```

## Introduction

This is the peer-graded course project of ["Getting and Cleaning Data"](https://www.coursera.org/learn/data-cleaning?specialization=data-science-foundations-r) on [Coursera](https://www.coursera.org/)

## Included Dataset

[Human Activity Recognition Using Smartphones Data Set](http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones)

The experiments have been carried out with a group of 30 volunteers within an age bracket of 19-48 years. Each person performed six activities (WALKING, WALKING_UPSTAIRS, WALKING_DOWNSTAIRS, SITTING, STANDING, LAYING) wearing a smartphone (Samsung Galaxy S II) on the waist. Using its embedded accelerometer and gyroscope, we captured 3-axial linear acceleration and 3-axial angular velocity at a constant rate of 50Hz. The experiments have been video-recorded to label the data manually. The obtained dataset has been randomly partitioned into two sets, where 70% of the volunteers was selected for generating the training data and 30% the test data.

The sensor signals (accelerometer and gyroscope) were pre-processed by applying noise filters and then sampled in fixed-width sliding windows of 2.56 sec and 50% overlap (128 readings/window). The sensor acceleration signal, which has gravitational and body motion components, was separated using a Butterworth low-pass filter into body acceleration and gravity. The gravitational force is assumed to have only low frequency components, therefore a filter with 0.3 Hz cutoff frequency was used. From each window, a vector of features was obtained by calculating variables from the time and frequency domain.


## Files Available
`Getting and Cleaning Data Course Project.Rproj`: r project file
`run_analysis.r`: code used to get and clean data  
`CodeBook`: description on variables, data and explanation of transformation performed to clean up the data  
`Tidy_data.csv`: tidy data output of `run_analysis.r`  
`UCI HAR Dataset/`: the raw dataset  
`README.md`

## Analysis Procedure
Let's review [Hadley Wickham's principles on tidy data](http://vita.had.co.nz/papers/tidy-data.pdf) first.

1. Each variable forms a column.
2. Each observation forms a row.
3. Each type of observational unit forms a table.

And [Hadley Wickham's standard on messy data](http://vita.had.co.nz/papers/tidy-data.pdf)

* Column headers are values, not variable names.  
* Multiple variables are stored in one column.  
* Variables are stored in both rows and columns.  
* Multiple types of observational units are stored in the same table. 
* A single observational unit is stored in multiple tables.

### Load Needed Packages and Files
```{r}
library(dplyr)
library(tidyr)

## Download and unzip the dataset if necessary
if (!dir.exists("UCI HAR Dataset")) {
        if (!file.exists("getdata-projectfiles-UCI HAR Dataset.zip")) {
                url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
                download.file(url, "./getdata-projectfiles-UCI HAR Dataset.zip", method = "curl")
        }
        unzip("getdata-projectfiles-UCI HAR Dataset.zip")
}


# Load the train and the test sets
activity_labels <- read.table("./UCI HAR Dataset/activity_labels.txt")[,2]
activity_labels <- tolower(activity_labels)
features <- read.table("./UCI HAR Dataset/features.txt")[,2]

subject_test <- read.table("./UCI HAR Dataset/test/subject_test.txt")
label_test <- read.table("./UCI HAR Dataset/test/Y_test.txt")
set_test <- read.table("./UCI HAR Dataset/test/X_test.txt")


subject_train <- read.table("./UCI HAR Dataset/train/subject_train.txt")
label_train <- read.table("./UCI HAR Dataset/train/Y_train.txt")
set_train <- read.table("./UCI HAR Dataset/train/X_train.txt")

```

### Merge Dataset
Combine variables containing subject, activity and measurement data into `activities`
```{r}
## Merge the train and the test sets to create one data set
subject_test <- mutate(subject_test, usage ="test")
subject_train <- mutate(subject_train, usage ="train")

test <- bind_cols(subject_test, label_test, set_test)
train <- bind_cols(subject_train, label_train, set_train)

activities <- bind_rows(test, train)
names(activities) <- c("subject", "usage", "activity", features)

activities <- as_tibble(activities, .name_repair = "minimal")
```

### Extract Only The Mean and Std Subset

Use `select()` for mean/std data screening
```{r}
activities <- activities %>%
        select(1:3, contains("mean()") | contains("std()")) %>%
        arrange(subject, activity)
```

### Add Specifier to Pivot

For each subject, the same activity is measured more than once, which makes subject id is not a unique observation specifier(or the `pivot_longer()` and `pivot_wider()` cannot match each observation's subject, mean and std correctly) for further cleaning, I add a specifier column manually.

In addtion, `pivot_longer()` and `pivot_wider()` are more developed version of outdated `gather()` and `spread()`.
```{r}
activities <- bind_cols(specifier = 1:nrow(activities), activities) %>%
        pivot_longer(cols = 5:70, names_to = "measurement", values_to = "value") 

```

### Replace `activities$activity` from Numbers to Characters
```{r}
activities$activity <- activity_labels[activities$activity]
activities$activity <- factor(activities$activity, levels = activity_labels)

activities$usage <- factor(activities$usage)
```

### Further Manipulation

Add a column to store what statistical method is used in `measurement` column, which will be the "names_from" parameters in next part' `pivot_wider()` function.
```{r}
data.mean <- activities[grep("mean", activities$measurement),] %>%
        mutate(method="mean")       

data.std <- activities[grep("std", activities$measurement),] %>%
        mutate(method="std") 

activities <- bind_rows(data.mean, data.std)

```

### Pivot Wider
Simplyfy the format of `measurement` in order to make `pivot_wider` work better.
Reshape the structure, make mean/std become separated colunms filled with corresponding value.
```{r}
activities$measurement <- sub("-(mean|std)[(][)]", "", activities$measurement)

activities <- pivot_wider(activities, names_from = method) %>%
        arrange(subject, activity, measurement)

activities$measurement <- factor(activities$measurement)

```

### Output Result

```{r}
write.csv(activities, "Tidy_data.csv")
```

